Binarity's Biases on Transit Survey Occurrence Rates

### 2018-02-28 paper/main.tex
remove appendix; change NPPS label; add brokenaxes
Better to not have Lambda(r) since we want discrete labels.

WRT the appendix:
As it was, the calculation that it's doing is roughly the same as what's
done in the main text.  However it tries to be more formal, and in
expanding the definition of "rate density" makes a mistake.

Defining Gamma(r,M) as number of planets per star per unit planet radius
per unit stellar mass is not actually the density function that we want.
This definition of Gamma is ill-defined because of the added measure in
stellar counts (the denominator). This problem is i

\title{ Binarity's Biases on Transit Survey Occurrence Rates }
%
\correspondingauthor{L. Bouma}
\email{luke@astro.princeton.edu}
%
\author[0000-0002-0514-5538]{L. G. Bouma}
\affiliation{
    Department of Astrophysical Sciences,
    Princeton University,
    4 Ivy Lane, Princeton, NJ 08540, USA}
\author[0000-0003-1298-9699]{K. Masuda}
\affiliation{
    Department of Astrophysical Sciences,
    Princeton University,
    4 Ivy Lane, Princeton, NJ 08540, USA}
\author[0000-0002-4265-047X]{J. N. Winn}
\affiliation{
    Department of Astrophysical Sciences,
    Princeton University,
    4 Ivy Lane, Princeton, NJ 08540, USA}
%
%
\begin{abstract}
%
Wide-field surveys for transiting planets, such as the NASA {\it
  Kepler} and {\it TESS} missions, are usually conducted without
knowing which stars have binary companions. Unresolved and
unrecognized binaries give rise to systematic errors in planet
occurrence rates, including misclassified planets and mistakes in
completeness corrections.  The individual errors can have different
signs, making it difficult to anticipate the net effect on inferred
occurrence rates. Here we use simplified models of signal-to-noise
limited transit surveys to try and clarify the situation. We derive a
formula for the apparent occurrence rate density measured by an
observer who falsely assumes all stars are single. The formula depends
on the binary fraction; the mass function of the secondary stars; and
the true occurrence of planets around primaries, secondaries, and
single stars. It also takes into account the Malmquist bias by which
binaries are over-represented in flux-limited samples. Application of
the formula to an idealized {\it Kepler}-like survey shows that for
planets larger than 2\,$R_\oplus$, the net systematic error is of
order 5\%.  In particular, unrecognized binaries are unlikely to be
the reason for the apparent discrepancies between hot Jupiter
occurrence rates measured in different surveys.  For smaller planets
the errors are potentially larger: the occurrence of Earth-sized
planets could be overestimated by as much as 50\%. We also show that
whenever high-resolution imaging reveals a transit host star to be a
binary, the planet is usually more likely to orbit the primary star
than the secondary star.
%
\end{abstract}
%
\keywords{
    methods: data analysis ---
    planets and satellites: detection ---
    surveys}
%
%

@@

Since the effects of binarity are most pronounced when the two stellar
components are similar, we begin by considering a universe in which
all stars are identical, with mass $M$, radius $R$, and luminosity
$L$.  Single stars are uniformly distributed in space with a number
density of $n\s$ stars per cubic parsec, and binary systems are
uniformly distributed with number density $n\b$.  In this scenario,
stars are never misclassified because the combined light of a binary
has the same color and spectrum as a single star.  We further assume
that all planets have the same radius, $r$, and occur around single
stars and members of binaries at the same rate, $\Lambda(r)$.

Our naive observers conduct a transit survey. To calculate the
occurrence rate of planets with radius $r$, they count the number of
detections of signals with amplitude $(r/R)^2$.  Then they identify
all the sources that appear to have been searchable for a signal of
amplitude $(r/R)^2$.  Here and throughout the rest of this paper, we
assume that the limiting source of noise is the photon-counting noise
from the source, i.e., $\sigma \propto 1/\sqrt{F}$, where
where $F$ is the total flux of the source.
Thus the observers determine the minimum $F_0$ for which detection would have been
possible, and count the number of sources with $F>F_0$. This will
include all the single stars out to a maximum distance
\begin{equation}
  d_0 = \sqrt{\frac{L}{4\pi F_0}}.
  \label{eq:dmax}
\end{equation}
Since binaries are twice as luminous, the condition $F>F_0$ will
include binaries out to the larger distance of $d_0\sqrt{2}$.
None of the stars in binaries will
appear to have a planet of radius $r$, because of the dilution of
the transit signal.  Thus, the {\it apparent} occurrence rate $\Lambda\a$ of
planets of radius $r$ is
\begin{equation}
    \Lambda\a(r) = 
        \frac{\Lambda(r) \, n\s d_0^3}
        {n\s d_0^3 + n\b (d_0\sqrt{2})^3}.
    \label{eq:rate_at_rp}
\end{equation}
This apparent rate is larger than the true rate by a factor
\begin{equation}
    \frac{\Lambda\a(r)}{\Lambda(r)} = 
        \frac{1}{1 + 2^{3/2}(n\b/n\s)}.
    \label{eq:correction_rp}
\end{equation}

The observers will also detect some transiting planets around stars
with binary companions.  The amplitude of these signals is
\begin{equation}
  \frac{L(r/R)^2}{L + L} = \frac{1}{2}\left(\frac{r}{R}\right)^2 =
  \left(\frac{r/\sqrt{2}}{R}\right)^2,
    \label{eq:delta_obs_general} 
\end{equation}
leading the astronomers to believe they have discovered a population
of planets with radius $r/\sqrt{2}$.  To calculate the corresponding
occurrence rate, they count the stars for which this type of signal
would have been detectable.  The limiting flux for detection in this
case is $4F_0$, because the signal amplitude has been reduced by a
factor of two and the noise level must also be reduced by a factor of
two.  The condition $F>4F_0$ is met for single stars within a distance
$d_0/2$, and binaries within a distance $d_0\sqrt{2}/2$.  Therefore,
the observers will calculate the occurrence rate of this new type of
planet to be
\begin{equation}
    \Lambda\a\left( \frac{r}{\sqrt{2}} \right) = 
        \frac{2\Lambda(r) \, n\b (d_0\sqrt{2}/2)^3}
        {n\s (d_0/2)^3 + n\b (d_0\sqrt{2}/2)^3}
    =
    \frac{2\Lambda(r)\cdot2^{3/2}(n\b/n\s)}{1 + 2^{3/2} (n\b/n\s)}.
    \label{eq:correction_diluted_rp}
\end{equation}
Figure~\ref{fig:model_1_volumes} illustrates the
volumes enclosing the apparently searchable single and binary stars.

@@

\begin{figure}[!tb]
    \begin{center}
        \includegraphics[width=.6\textwidth]{figures/occ_rate_vs_radius_model_1_brokenx.pdf}
    \end{center}
    \vspace{-0.5cm}
    \caption{ Apparent occurrence rate $\Lambda\a$, and true
    occurrence rate $\Lambda$ in a universe with only one type of
    star, one type of planet (with radius $R_\oplus$), and 
    a binary system number density 5\% that of single stars.
    The occurrence
    rate of planets with radius $R_\oplus$ is underestimated, and the
    occurrence rate of planets with radius
    $R_\oplus/\sqrt{2}$ is overestimated.}
    \label{fig:occ_rate_model_1}
\end{figure}

We can now assess the severity of the errors, for a given value of the
binary-to-single ratio $n\b/n\s$. For stars with masses from 0.7 to
1.3~$M_\odot$, \citet{raghavan_survey_2010} found the multiplicity
fraction~--~the fraction of systems in a volume-limited sample that
are multiple~--~to be 0.44.  Assuming all multiple systems are
binaries, this gives a binary fraction
\begin{equation}
  \frac{n\b}{n\s + n\b} \approx 0.44,
\end{equation}
which implies $n\b/n\s \approx 0.79$.  Of course not all of
these binaries are ``twin'' binaries as we have assumed in our simple
calculation.  Later, in Section~\ref{sec:general_formula}, we will
allow for a continuum of properties for the secondary stars. For now,
we might guess that only a tenth of the binaries have pairs of stars
close enough in brightness to produce errors as significant as those
we have been considering.  Adopting the value $n\b/n\s\approx0.05$, we
find
\begin{equation}
    \frac{\Lambda\a(r)}{\Lambda(r)} = 0.88,~~
    \frac{\Lambda\a(r/\sqrt{2})}{\Lambda(r)} = 0.25.
    \label{eq:ratios}
\end{equation}
All together, the various effects produce biases of order 10\% in the
occurrence rates.  We will see that this level of error is
characteristic of many of our more complicated models as well.

@@

We will now generalize to allow for a continuum of planet sizes.   We
introduce the occurrence rate density
\begin{equation}
    \Gamma(r) \equiv \frac{{\rm d}\Lambda}{{\rm d}r},
    \label{eq:rate_density_defn}
\end{equation}
the number of planets per star per unit planet radius.  The naive
observers are measuring $\Gamma\a(r\a)$, the apparent occurrence rate
density of planets with an ``apparent radius''
\begin{equation}
  r\a = R \sqrt{\delta}.
\end{equation}

@@

From within this sample of apparently searchable sources, the
observers sum the number of detections ${\rm d}N_{\rm det}$ of planets
with an apparent radius between $r\a$ and $r\a + dr\a$.  There will be
three contributions to this sum,
\begin{align}
  {\rm d}N_{{\rm det},0} &= 
      \frac{N_\star(r\a) p_{\rm tra}}{1+\mu}
      \Gamma_0(r\a)\mathrm{d}r\a,
  \label{eq:n0} \\
  {\rm d}N_{{\rm det},1} &= 
      \frac{\mu N_\star(r\a) p_{\rm tra}}{1+\mu} 
      \Gamma_1(\sqrt{2}r\a)\mathrm{d}(\sqrt{2}r\a),
	\label{eq:n1} \\
  {\rm d}N_{{\rm det},2} &=
      \frac{\mu N_\star(r\a) p_{\rm tra}}{1+\mu} 
      \Gamma_2(\sqrt{2}r\a)\mathrm{d}(\sqrt{2}r\a).
	\label{eq:n2}
\end{align}
where here and elsewhere, the subscript 0 refers to single stars, 1
refers to primary stars of binaries, and 2 refers to secondary stars
of binaries. The apparent occurrence rate density is therefore
\begin{equation}
  \Gamma\a(r\a) = \frac{{\rm d}N_{\rm det}}{{\rm d}r\a} \frac{1}{N_\star(r\a) p_{\rm tra}} =
  \frac{\Gamma_0(r\a)}{1+\mu} +
  \frac{\mu\sqrt{2}\,\Gamma_1(\sqrt{2}r\a)}{1+\mu} +
  \frac{\mu\sqrt{2}\,\Gamma_2(\sqrt{2}r\a)}{1+\mu}.
	\label{eq:formula_twin_binary}
\end{equation}
Our earlier results, Equations~\ref{eq:rate_at_rp}
and~\ref{eq:correction_diluted_rp}, are recovered by integrating this formula
over $r$ after inserting
\begin{equation}
  \Gamma_i(r) = \Lambda(r\p)\, \hat{\delta}(r-r\p),~~{\rm for}\ i\in\{0,1,2\}
\end{equation}
where $\hat{\delta}$ is the Dirac delta function, and $r\p$ is the
single planet size considered in Section~\ref{sec:model_1}.

@@

To compute the apparent occurrence rate density, we need to make
the following modifications to Equation~\ref{eq:formula_twin_binary}:
%
\begin{enumerate}
%
\item The Malmquist bias is different.  For a given value of $r\a$, the
  number of binary systems in the searchable sample with mass ratio
  $(q, q+\mathrm{d}q)$ is
\begin{equation}
    {N_\star(r\a)\over{1+\mu}}\, {n_{\rm b}\over n_{\rm s}}
    \left[{L(M) + L(qM) \over L(M)}\right]^{3/2}
    f(q)\,\mathrm{d}q,
    \label{eq:N_binaries}
\end{equation}
where $\mu$ is given by
\begin{equation}
    \mu= 
    %\frac{N_\star^{\rm b}(r\a)}{N_s^0(r\a)}=
    \int_0^1 {n_{\rm b}\over n_{\rm s}}\left[{L(M) + L(qM)
    \over L(M\a)}\right]^{3/2} f(q)\,\mathrm{d}q.
    \label{eq:mu_general}
\end{equation}
To obtain the total number of searchable binaries,
Equation~\ref{eq:N_binaries} needs to be integrated over $q$. The
result of the integration replaces ${\mu N_\star(r\a)/(1+\mu)}$ in
Equations~\ref{eq:n1} and~\ref{eq:n2}.

@@

Taking these modifications into account, a general formula for the
apparent rate density is
\begin{align}
    \notag
    \Gamma\a(r\a) &= \frac{1}{1+\mu} \,
    \left\{ \Gamma_0(r\a)+ 
    \frac{n\b}{n\s}
    \left[ \int_0^1 \mathrm{d}q \,
           \mathcal{D}_1^3 f(q)\cdot
           \mathcal{D}_1 \Gamma_1(\mathcal{D}_1r\a)\,
    \right.   
    \right. \\
    & \quad\quad\quad\quad\quad \left.\left.
    +\int_0^1 {\rm d}q\, 
         \mathcal{D}_1^3 f(q)\cdot \mathcal{D}_2
         \Gamma_2(\mathcal{D}_2r\a)\cdot
         \frac{R(qM)}{R(M)} q^{-1/3}
    \right] \right\}.
    \label{eq:general_Gamma_a}
\end{align}
The \hyperref[sec:appendix]{appendix} gives an alternative derivation
of this equation.

We note that the combination $\mathcal{D}_1^3f(q)$ is the mass-ratio
distribution for binaries contained within the searchable sample of
sources.  With this in mind we may rewrite
Equation~\ref{eq:general_Gamma_a} as
\begin{align}
    %\notag
    %\notag
    \Gamma\a(r\a)
    =%&= 
    {1\over 1+\mu} \,
    \left[
       \Gamma_0(r\a)%\phantom{\tilde \Gamma}\right.%&\\
       %&\left.
       +\mu\,
       \left\langle
       \mathcal{D}_1\cdot\Gamma_1\left(\mathcal{D}_1r\a\right)
       +
       {q \mathcal{D}_2}\cdot\Gamma_2\left(\mathcal{D}_2r\a\right)
       \cdot
       {R(qM) \over R(M)}q^{-1/3}
       \right\rangle
    \right],
\end{align}
where the angle brackets denote averaging over all the binaries in the
searchable sample.

@@

\begin{figure}[!t]
    \centering
    \epsscale{1.15}
    \plottwo{figures/int_rate_density_vs_radius_model_3_rpu_22.5_manyZs.pdf}{figures/int_occ_rate_vs_radius_model_3_rpu_22.5_manyZs.pdf}
    \caption{ {\it Left.}---Apparent occurrence rate density
      ($\Gamma\a$), compared to the true occurrence rate density
      ($\Gamma_0$) for single stars and primary stars in binaries.
      Three different cases are plotted, with different choices for
      the number of planets per secondary star relative to
      that for single stars ($Z_2/Z_0$).
      {\it Right.}---Same, but for apparent occurrence
      rate ($\Lambda\a$) within $0.5\,R_\oplus$ bins.
      In both cases the true planet
      radius distribution is specified by
      Equation~\ref{eq:model3_radius_distribution}.
    }
    \label{fig:occ_rate_model_3_log}
\end{figure}

For planets smaller than 2\,$R_\oplus$, the true occurrence rate
density is more uncertain because such planets are more difficult to
detect.  To investigate the systematic errors associated with binarity
on the inferred occurrence rate of Earth-sized planets, we need to
make plausible assumptions about the true occurrence rate density.
Obviously if the rate density continues to vary as $r^{-3}$ down to
much smaller planet sizes, the results of the preceding section will
hold.  But there is no particular reason to think this will be the
case and indeed, some investigators have concluded that the rate
density begins to level off to a constant value as $r$ decreases below
$R_\oplus$~\citep{petigura_plateau_2013}.  To investigate the
implications we consider a broken power-law:
\begin{align}
    f(r)
    &\propto
    \left.
    \begin{cases}
        r^\gamma & \text{for } r\geq 2R_\oplus \\
        {\rm constant} & \text{for } r\leq2R_\oplus.
    \end{cases}
    \right.
    \label{eq:model3_radius_distribution}
\end{align}
In this case the integrals that appear in
Equation~\ref{eq:general_Gamma_a} are tedious to work out
analytically, leading us to evaluate them numerically.\footnote{We
refer the interested reader to our online code for performing this
computation: \url{github.com/lgbouma/binary_biases}, commit
\texttt{4e517fe}.} Figure~\ref{fig:occ_rate_model_3_log} shows the
results, again for the case $n\b/n\s=0.79$, $\alpha = 3.5$, $\beta=0$,
and $\gamma=-2.92$.

@@

Our formalism also provides a way to calculate the relative
probability that a detected planet orbits a single star, the primary
star, or a secondary star.  These can be obtained from the
generalizations of Equations~\ref{eq:n0}--\ref{eq:n2}, which are
derived in the \hyperref[sec:appendix]{appendix}.  This information
can be used to help interpret the results of a transit survey.  It is
also relevant to the fairly common situation in which transits are
detected from a source, and subsequent high-resolution imaging reveals
the source to be a multiple-star system.  The question then arises:
which star is the host of the detected planet?  Without further
observations the answer is often unclear.  To illustrate we return to
the scenario described in the previous section: a broken power-law
radius distribution with $\gamma=-2.92$, along with the choices
$n\b/n\s=0.79$, $\alpha = 3.5$, and $\beta=0$.  We also try different
assumptions for the rate of planets around secondaries, i.e.,
different choices for $Z_2/Z_0$.  We use
Equations~\ref{eq:n_det_0}--\ref{eq:n_det_2} to compute the number of
detections from single stars, primaries, and secondaries, and divide
by the total number of detections at each apparent radius.

@@

\acknowledgements{
It was a pleasure discussing this study with T.~Barclay, W.~Bhatti,
J.~Christiansen, F.~Dai, and T.~Morton.  This work made use of NASA's
Astrophysics Data System Bibliographic Services.
%
\newline
%
\software{\texttt{numpy}~\citep{walt_numpy_2011}, 
\texttt{scipy}~\citep{jones_scipy_2001}, 
\texttt{matplotlib}~\citep{hunter_matplotlib_2007}, 
\texttt{pandas}~\citep{mckinney-proc-scipy-2010},
\texttt{IPython}~\citep{perez_2007}
}
}

\newpage
\appendix
\section{Alternative derivation of apparent rate density}
\label{sec:appendix}

>>>>

\title{ Binarity's Biases on Transit Survey Occurrence Rates }
%
\correspondingauthor{L. Bouma}
\email{luke@astro.princeton.edu}
%
\author[0000-0002-0514-5538]{L. G. Bouma}
\affiliation{
    Department of Astrophysical Sciences,
    Princeton University,
    4 Ivy Lane, Princeton, NJ 08540, USA}
\author[0000-0003-1298-9699]{K. Masuda}
\affiliation{
    Department of Astrophysical Sciences,
    Princeton University,
    4 Ivy Lane, Princeton, NJ 08540, USA}
\affiliation{
    NASA Sagan Fellow}
\author[0000-0002-4265-047X]{J. N. Winn}
\affiliation{
    Department of Astrophysical Sciences,
    Princeton University,
    4 Ivy Lane, Princeton, NJ 08540, USA}
%
%
\begin{abstract}
%
Wide-field surveys for transiting planets, such as the NASA {\it
  Kepler} and {\it TESS} missions, are usually conducted without
knowing which stars have binary companions. Unresolved and
unrecognized binaries give rise to systematic errors in planet
occurrence rates, including misclassified planets and mistakes in
completeness corrections.  The individual errors can have different
signs, making it difficult to anticipate the net effect on inferred
occurrence rates. Here we use simplified models of signal-to-noise
limited transit surveys to try and clarify the situation. We derive a
formula for the apparent occurrence rate density measured by an
observer who falsely assumes all stars are single. The formula depends
on the binary fraction; the mass function of the secondary stars; and
the true occurrence of planets around primaries, secondaries, and
single stars. It also takes into account the Malmquist bias by which
binaries are over-represented in flux-limited samples. Application of
the formula to an idealized {\it Kepler}-like survey shows that for
planets larger than 2\,$R_\oplus$, the net systematic error is of
order 5\%.  In particular, unrecognized binaries are unlikely to be
the reason for the apparent discrepancies between hot Jupiter
occurrence rates measured in different surveys.  For smaller planets
the errors are potentially larger: the occurrence of Earth-sized
planets could be overestimated by as much as 50\%. We also show that
whenever high-resolution imaging reveals a transit host star to be a
binary, the planet is usually more likely to orbit the primary star
than the secondary star.
%
\end{abstract}
%
\keywords{
    methods: data analysis ---
    planets and satellites: detection ---
    surveys}
%
%

@@

Since the effects of binarity are most pronounced when the two stellar
components are similar, we begin by considering a universe in which
all stars are identical, with mass $M$, radius $R$, and luminosity
$L$.  Single stars are uniformly distributed in space with a number
density of $n\s$ stars per cubic parsec, and binary systems are
uniformly distributed with number density $n\b$.  In this scenario,
stars are never misclassified because the combined light of a binary
has the same color and spectrum as a single star.  We further assume
that all planets have the same radius, $r$, and occur around single
stars and members of binaries at the same rate, $\Lambda_r$.

Our naive observers conduct a transit survey. To calculate the
occurrence rate of planets with radius $r$, they count the number of
detections of signals with amplitude $(r/R)^2$.  Then they identify
all the sources that appear to have been searchable for a signal of
amplitude $(r/R)^2$.  Here and throughout the rest of this paper, we
assume that the limiting source of noise is the photon-counting noise
from the source, i.e., $\sigma \propto 1/\sqrt{F}$,
where $F$ is the total flux of the source.
Thus the observers determine the minimum $F_0$ for which detection would have been
possible, and count the number of sources with $F>F_0$. This will
include all the single stars out to a maximum distance
\begin{equation}
  d_0 = \sqrt{\frac{L}{4\pi F_0}}.
  \label{eq:dmax}
\end{equation}
Since binaries are twice as luminous, the condition $F>F_0$ will
include binaries out to the larger distance of $d_0\sqrt{2}$.
None of the stars in binaries will
appear to have a planet of radius $r$, because of the dilution of
the transit signal.  Thus, the {\it apparent} occurrence rate $\Lambda\a$ of
planets of radius $r$ is
\begin{equation}
    \Lambda_{{\rm{a}},r} = 
        \frac{\Lambda_r \, n\s d_0^3}
        {n\s d_0^3 + n\b (d_0\sqrt{2})^3}.
    \label{eq:rate_at_rp}
\end{equation}
This apparent rate is smaller than the true rate by a factor
\begin{equation}
    \frac{\Lambda_{{\rm{a}},r}}{\Lambda_r} = 
        \frac{1}{1 + 2^{3/2}(n\b/n\s)}.
    \label{eq:correction_rp}
\end{equation}

The observers will also detect some transiting planets around stars
with binary companions.  The amplitude of these signals is
\begin{equation}
  \frac{L(r/R)^2}{L + L} = \frac{1}{2}\left(\frac{r}{R}\right)^2 =
  \left(\frac{r/\sqrt{2}}{R}\right)^2,
    \label{eq:delta_obs_general} 
\end{equation}
leading the astronomers to believe they have discovered a population
of planets with radius $r/\sqrt{2}$.  To calculate the corresponding
occurrence rate, they count the stars for which this type of signal
would have been detectable.  The limiting flux for detection in this
case is $4F_0$, because the signal amplitude has been reduced by a
factor of two and the noise level must also be reduced by a factor of
two.  The condition $F>4F_0$ is met for single stars within a distance
$d_0/2$, and binaries within a distance $d_0\sqrt{2}/2$.  Therefore,
the observers will calculate the occurrence rate of this new type of
planet to be
\begin{equation}
    \Lambda_{{\rm{a}},r/\sqrt{2}} = 
        \frac{2\, \Lambda_r \, n\b (d_0\sqrt{2}/2)^3}
        {n\s (d_0/2)^3 + n\b (d_0\sqrt{2}/2)^3}
    =
    \frac{2\, \Lambda_r \cdot2^{3/2}(n\b/n\s)}{1 + 2^{3/2} (n\b/n\s)}.
    \label{eq:correction_diluted_rp}
\end{equation}
Figure~\ref{fig:model_1_volumes} illustrates the
volumes enclosing the apparently searchable single and binary stars.

@@

\begin{figure}[!tb]
    \begin{center}
        \includegraphics[width=.6\textwidth]{figures/occ_rate_vs_radius_model_1_brokenx.pdf}
    \end{center}
    \vspace{-0.5cm}
    \caption{ Apparent occurrence rate $\Lambda\a$, and true
    occurrence rate $\Lambda$ in a universe with only one type of
    star, one type of planet (with radius $r$), and
    $n\b/n\s=0.05$, for $n\b$ the number density of binary
    systems, and $n\s$ the number density of single stars.
    The occurrence
    rate of planets with radius $r$ is underestimated, and the
    occurrence rate of planets with radius
    $r/\sqrt{2}$ is overestimated.}
    \label{fig:occ_rate_model_1}
\end{figure}

We can now assess the severity of the errors, for a given value of the
binary-to-single ratio $n\b/n\s$. For stars with masses from 0.7 to
1.3~$M_\odot$, \citet{raghavan_survey_2010} found the multiplicity
fraction~--~the fraction of systems in a volume-limited sample that
are multiple~--~to be 0.44.  Assuming all multiple systems are
binaries, this gives a binary fraction
\begin{equation}
  \frac{n\b}{n\s + n\b} \approx 0.44,
\end{equation}
which implies $n\b/n\s \approx 0.79$.  Of course not all of
these binaries are ``twin'' binaries as we have assumed in our simple
calculation.  Later, in Section~\ref{sec:general_formula}, we will
allow for a continuum of properties for the secondary stars. For now,
we might guess that only a tenth of the binaries have pairs of stars
close enough in brightness to produce errors as significant as those
we have been considering.  Adopting the value $n\b/n\s\approx0.05$, we
find
\begin{equation}
    \frac{\Lambda_{{\rm{a}},r}}{\Lambda_r} = 0.88,~~
    \frac{\Lambda_{{\rm{a}},r/\sqrt{2}}}{\Lambda_r} = 0.25.
    \label{eq:ratios}
\end{equation}
All together, the various effects produce biases of order 10\% in the
occurrence rates.  We will see that this level of error is
characteristic of many of our more complicated models as well.

@@

We will now generalize to allow for a continuum of planet sizes.   We
introduce the occurrence rate density $\Gamma$,
the number of planets per star per unit planet radius, so that
\begin{equation}
    \Gamma(r) {\rm d}r \equiv \frac{ {\rm d}N_{\rm det} }{N_\star}
            \, \frac{1}{p_{\rm tra}}.
    \label{eq:rate_density_defn}
\end{equation}
The naive observers are measuring $\Gamma\a(r\a)$, the apparent
occurrence rate density of planets with an ``apparent radius''
\begin{equation}
  r\a = R \sqrt{\delta}.
\end{equation}

@@

From within this sample of apparently searchable sources, the
observers sum the number of detections ${\rm d}N_{\rm det}$ of planets
with an apparent radius between $r\a$ and $r\a + dr\a$.  There will be
three contributions to this sum,
\begin{align}
  {\rm d}N_{{\rm det},0} &= 
      \frac{N_\star(r\a) p_{\rm tra}}{1+\mu}
      \Gamma_0(r\a)\mathrm{d}r\a,
  \label{eq:n0} \\
  {\rm d}N_{{\rm det},1} &= 
      \frac{\mu N_\star(r\a) p_{\rm tra}}{1+\mu} 
      \Gamma_1(\sqrt{2}r\a)\mathrm{d}(\sqrt{2}r\a),
	\label{eq:n1} \\
  {\rm d}N_{{\rm det},2} &=
      \frac{\mu N_\star(r\a) p_{\rm tra}}{1+\mu} 
      \Gamma_2(\sqrt{2}r\a)\mathrm{d}(\sqrt{2}r\a).
	\label{eq:n2}
\end{align}
where here and elsewhere, the subscript 0 refers to single stars, 1
refers to primary stars of binaries, and 2 refers to secondary stars
of binaries. The apparent occurrence rate density is therefore
\begin{equation}
  \Gamma\a(r\a) = \frac{{\rm d}N_{\rm det}}{{\rm d}r\a} \frac{1}{N_\star(r\a) p_{\rm tra}} =
  \frac{\Gamma_0(r\a)}{1+\mu} +
  \frac{\mu\sqrt{2}\,\Gamma_1(\sqrt{2}r\a)}{1+\mu} +
  \frac{\mu\sqrt{2}\,\Gamma_2(\sqrt{2}r\a)}{1+\mu}.
	\label{eq:formula_twin_binary}
\end{equation}
Our earlier results, Equations~\ref{eq:rate_at_rp}
and~\ref{eq:correction_diluted_rp}, are recovered by integrating this formula
over $r$ after inserting
\begin{equation}
  \Gamma_i(r) = \Lambda_{r\p}\, \hat{\delta}(r-r\p),~~{\rm for}\ i\in\{0,1,2\}
\end{equation}
where $\hat{\delta}$ is the Dirac delta function, and $\Lambda_{r\p}$ is the
number of planets with size $r\p$ per star considered in 
Section~\ref{sec:model_1}.

@@

To compute the apparent occurrence rate density, we need to make
the following modifications to Equation~\ref{eq:formula_twin_binary}:
%
\begin{enumerate}
%
\item The Malmquist bias is different.  For a given value of $r\a$, the
  number of binary systems in the searchable sample with mass ratio
  $(q, q+\mathrm{d}q)$ is
\begin{equation}
    {N_\star(r\a)\over{1+\mu}}\, {n_{\rm b}\over n_{\rm s}}
    \left[{L(M) + L(qM) \over L(M)}\right]^{3/2}
    f(q)\,\mathrm{d}q,
    \label{eq:N_binaries}
\end{equation}
where $\mu$ is given by
\begin{equation}
    \mu= 
    %\frac{N_\star^{\rm b}(r\a)}{N_s^0(r\a)}=
    \int_0^1 {n_{\rm b}\over n_{\rm s}}\left[{L(M) + L(qM)
    \over L(M\a)}\right]^{3/2} f(q)\,\mathrm{d}q.
    \label{eq:mu_general}
\end{equation}
Equation~\ref{eq:N_binaries} replaces ${\mu N_\star(r\a)/(1+\mu)}$
in Equations~\ref{eq:n1} and~\ref{eq:n2}.
The whole equation is then integrated to give the number of
detected planets.

@@

Taking these modifications into account, a general formula for the
apparent rate density is
\begin{align}
    \notag
    \Gamma\a(r\a) &= \frac{1}{1+\mu} \,
    \left\{ \Gamma_0(r\a)+ 
    \frac{n\b}{n\s}
    \left[ \int_0^1 \mathrm{d}q \,
           \mathcal{D}_1^3 f(q)\cdot
           \mathcal{D}_1 \Gamma_1(\mathcal{D}_1r\a)\,
    \right.   
    \right. \\
    & \quad\quad\quad\quad\quad \left.\left.
    +\int_0^1 {\rm d}q\, 
         \mathcal{D}_1^3 f(q)\cdot \mathcal{D}_2
         \Gamma_2(\mathcal{D}_2r\a)\cdot
         \frac{R(qM)}{R(M)} q^{-1/3}
    \right] \right\}.
    \label{eq:general_Gamma_a}
\end{align}
%The \hyperref[sec:appendix]{appendix} gives a more formal derivation
%of this equation.
We note that the combination $\mathcal{D}_1^3f(q)$ is the mass-ratio
distribution for binaries contained within the searchable sample of
sources.  With this in mind we may rewrite
Equation~\ref{eq:general_Gamma_a} as
\begin{align}
    %\notag
    %\notag
    \Gamma\a(r\a)
    =%&= 
    {1\over 1+\mu} \,
    \left[
       \Gamma_0(r\a)%\phantom{\tilde \Gamma}\right.%&\\
       %&\left.
       +\mu\,
       \left\langle
       \mathcal{D}_1\cdot\Gamma_1\left(\mathcal{D}_1r\a\right)
       +
       {\mathcal{D}_2}\cdot\Gamma_2\left(\mathcal{D}_2r\a\right)
       \cdot
       {R(qM) \over R(M)}q^{-1/3}
       \right\rangle
    \right],
\end{align}
where the angle brackets denote averaging over all the binaries in the
searchable sample.

@@

\begin{figure}[!t]
    \centering
    \epsscale{1.15}
    \plottwo{figures/int_rate_density_vs_radius_model_3_rpu_22.5_manyZs.pdf}{figures/int_occ_rate_vs_radius_model_3_rpu_22.5_manyZs.pdf}
    \caption{ {\it Left.}---Apparent occurrence rate density
      ($\Gamma\a$), compared to the true occurrence rate density
      ($\Gamma_0$) for single stars and primary stars in binaries.
      Three different cases are plotted, with different choices for
      the number of planets per secondary star relative to
      that for single stars ($Z_2/Z_0$).
      {\it Right.}---Same, integrated over $0.5\,R_\oplus$
      bins to give apparent and true occurrence rates.
      In both cases the true planet
      radius distribution is specified by
      Equation~\ref{eq:model3_radius_distribution}.
    }
    \label{fig:occ_rate_model_3_log}
\end{figure}

For planets smaller than 2\,$R_\oplus$, the true occurrence rate
density is more uncertain because such planets are more difficult to
detect.  To investigate the systematic errors associated with binarity
on the inferred occurrence rate of Earth-sized planets, we need to
make plausible assumptions about the true occurrence rate density.
Obviously if the rate density continues to vary as $r^{-3}$ down to
much smaller planet sizes, the results of the preceding section will
hold.  But there is no particular reason to think this will be the
case and indeed, some investigators have concluded that the rate
density begins to level off to a constant value as $r$ decreases below
$2R_\oplus$~\citep{petigura_plateau_2013}.  To investigate the
implications we consider a broken power-law:
\begin{align}
    f(r)
    &\propto
    \left.
    \begin{cases}
        r^\gamma & \text{for } r\geq 2R_\oplus \\
        {\rm constant} & \text{for } r\leq2R_\oplus.
    \end{cases}
    \right.
    \label{eq:model3_radius_distribution}
\end{align}
In this case the integrals that appear in
Equation~\ref{eq:general_Gamma_a} are tedious to work out
analytically, leading us to evaluate them numerically.\footnote{We
refer the interested reader to our online code for performing this
computation: \url{github.com/lgbouma/binary_biases}, commit
\texttt{4e517fe}.} Figure~\ref{fig:occ_rate_model_3_log} shows the
results, again for the case $n\b/n\s=0.79$, $\alpha = 3.5$, $\beta=0$,
and $\gamma=-2.92$.

@@

Our formalism also provides a way to calculate the relative
probability that a detected planet orbits a single star, the primary
star, or a secondary star.  
This information can be used to help interpret the results of
a transit survey.  It is
also relevant to the fairly common situation in which transits are
detected from a source, and subsequent high-resolution imaging reveals
the source to be a multiple-star system.  The question then arises:
which star is the host of the detected planet?  Without further
observations the answer is often unclear.  To illustrate we return to
the scenario described in the previous section: a broken power-law
radius distribution with $\gamma=-2.92$, along with the choices
$n\b/n\s=0.79$, $\alpha = 3.5$, and $\beta=0$.  We also try different
assumptions for the rate of planets around secondaries, i.e.,
different choices for $Z_2/Z_0$.  
Dividing each component of Equation~\ref{eq:general_Gamma_a} by
the total rate density, we get the fraction of
detections from single stars, primaries, and secondaries as a function
of apparent radius.

@@

\acknowledgements{
It was a pleasure discussing this study with T.~Barclay, W.~Bhatti,
J.~Christiansen, F.~Dai, and T.~Morton.  This work made use of NASA's
Astrophysics Data System Bibliographic Services.
This work was performed in part under contract with the California Institute 
of Technology (Caltech)/Jet Propulsion Laboratory (JPL) funded by NASA through 
the Sagan Fellowship Program executed by the NASA Exoplanet Science Institute.
%
\newline
%
\software{\texttt{numpy}~\citep{walt_numpy_2011}, 
\texttt{scipy}~\citep{jones_scipy_2001}, 
\texttt{matplotlib}~\citep{hunter_matplotlib_2007}, 
\texttt{pandas}~\citep{mckinney-proc-scipy-2010},
\texttt{IPython}~\citep{perez_2007}
}
}

\newpage

\begin{comment}
\appendix
\section{Alternative derivation of apparent rate density}
\label{sec:appendix}
